Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
added more to paper

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.75\linewidth]{parareal_visualization.png}
\caption{Graphical representation of the Parareal scheme.}
\label{fig:pararealvis}
\end{center}
\end{figure}

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.

@@

\begin{itemize}
\item General Scaling with Incrased $N$
\item Adjusting $K$ in the Parareal
\item Using Different Integration Schemes in Parareal for $g_{\textnormal{fine}}$
\item Run the Solver on Various Complexities of ODEs
\end{itemize}

@@

\subsection{General Remarks about Results}
LETS SEE IF I HAVE TO DISCLAIM ANY GLOBAL FAILURES IN OUR RESULTS

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
\caption{Running time vs. number of CPUs.}
\label{fig:run_v_cpu}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
\caption{Error vs. number of correction steps.}
\label{fig:err_v_k}
\end{center}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
\caption{Error vs. quality factor of the fine operator}
\label{fig:err_v_q}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
\caption{Running time as a function of the number of correction steps.}
\label{fig:run_v_k}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\linewidth]{decomp.png}
\caption{Parallelization by space schematic.}
\label{fig:spaceparallel}
\end{center}
\end{figure}

@@

\subsection{Summary of Performance}
\subsubsection{Parallel in Space Paradigm}
In serial, we have:

@@

\subsection{Demonstrating Scaling and Efficiency}
\subsubsection{Parallel in Space Paradigm}
Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved. However, scaling is \textbf{not} achieved for the 256$\times$ 256 grid when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed). We do see scaling with approximately $50\%$ efficiency for the $512\times 512$ grid as we add more processors. From the results, as we go from the $256\times256$ grid to the $512\times 512$ grid, we quadruple the size of the problem; quadrupling the processors from $2\times 2$ to $4\times4$ yields a somewhat linear relationship. Hence, this is an example of weak scaling. An advantage to this paradigm is as follows: because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes. Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.

@@

\section{Appendix}
\subsection{Explanation of Code Files}

>>>>

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\linewidth]{parareal_visualization.png}
\caption{Graphical representation of the Parareal scheme.}
\label{fig:pararealvis}
\end{center}
\end{figure}

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$ explained in the below table.

@@

\begin{itemize}
\item General Scaling with Increased $N$
\item Adjusting $K$ in the Parareal
\item Using Different Integration Schemes in Parareal for $g_{\textnormal{fine}}$
\item Run the Solver on Various Complexities of ODEs
\end{itemize}

@@

\subsection{General Remarks about Results}

In general, we were unable to see the promised theoretical speedups and error
convergence for the Parareal algorithm. We believe this could be because of
several factors. First, we may have some subtle bugs in our implementation of
the Parareal algorithms. Additionally, we only tested our implementation on
relatively simple problems.  he problem with this testing approach is that the
fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.
We did not see the running time of the Parareal beat that for the serial
implementation of $g_{\textnormal{fine}}$. If we tested our implementation for
more expensive $g_{\textnormal{fine}}$ and more complicated problems, we might
be able to see to see more speedup and better error convergence.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
\caption{Running time vs. number of CPUs.}
\label{fig:run_v_cpu}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
\caption{Error vs. number of correction steps.}
\label{fig:err_v_k}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
\caption{Error vs. quality factor of the fine operator}
\label{fig:err_v_q}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
\caption{Running time as a function of the number of correction steps.}
\label{fig:run_v_k}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.5\linewidth]{decomp.png}
\caption{Parallelization by space schematic.}
\label{fig:spaceparallel}
\end{center}
\end{figure}

@@

In serial, we have:

@@

Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved. However, scaling is \textbf{not} achieved for the 256$\times$ 256 grid when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed). We do see scaling with approximately $50\%$ efficiency for the $512\times 512$ grid as we add more processors. From the results, as we go from the $256\times256$ grid to the $512\times 512$ grid, we quadruple the size of the problem; quadrupling the processors from $2\times 2$ to $4\times4$ yields a somewhat linear relationship. Hence, this is an example of weak scaling. An advantage to this paradigm is as follows: because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes. Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.

@@
